\section{Statistical Mechanics of Generalization (\SMOG)}
\label{sxn:SMOG_main}

In this section, we review \STATMECH: 
both to understand how it is usually applied in \StatisticalMechanicsOfGeneralization (\SMOG) theory; 
and to understand how our \SemiEmpirical approach in \SETOL is similar to and different from the traditional approach.
The goal is to obtain an expression for the \GeneralizationAccuracy (or \ModelQuality $\Q^{ST}$) for the
classic \StudentTeacher (ST) model of the Linear \Perceptron (in the AA, and at high-T),
as described in \cite{SST90,SST92}.
In Section~\ref{sxn:matgen}, we will generalize this to a \LayerQuality metric, $\Q$, for a layer in a general Multi-Layer Perceptron (MLP),
i.e., $\Q^{ST}\rightarrow\Q$, so that $\Q$ can then be expressed in terms of the ESD of the NN layer.

\paragraph{Outline.}
Here is an outline of this section.
\begin{enumerate}[label=4.\arabic*]
\item
  \textbf{Traditional Approaches to the \SMOG}
  In Section~\ref{sxn:trad_smog}, we explain the mapping from the \STATMECH theory of disordered systems
    to the \STATMECH theory of NN learning (\SMOG); and, additionally, how our \SemiEmpirical approach
    is similar to and different from the traditional approach.

  \item
      \textbf{Mathematical Preliminaries}
    In Section~\ref{sxn:mathP}, we review the mathematical details of \STATMECH, providing definitions
      and detailed derivations of quantities and expressions necessary later.  

    \item
        \textbf{Student-Teacher Model:}
      In Section~\ref{sxn:SMOG_main-student_teacher}, we discuss the setup of the \StudentTeacher (ST) model
    as a general means to estimate the \AverageGeneralizationError empirically.  

  \item
      \textbf{Generalization Gap vs. Model Quality}
    In Section~\ref{sxn:SMOG_main-model_quality}, we briefly highlight the differences between the \STATMECH model
    of the \GeneralizationAccuracy and the ST model for the \GeneralizationGap.

  \item
      \textbf{Bad Training Data and the Spin Glass Phase}
    In Section~\ref{sxn:SMOG_main-spin_glass}, we describe the qualitative aspects of spin-glass theory
    and its implications for the ST model.

  \item
      \textbf{Average Generalization Error $(\AVGSTGE)$}
    In Section~\ref{sxn:SMOG_main-st_av}, we derive the (new) result for the
    ST \ModelQuality, $\Q^{ST}$, using the setup of the classic (ST) model for the
    \GeneralizationError (and accuracy) of the \Perceptron (in the AA, and at high-T).
\end{enumerate}

\noindent
Additional information can be found in the Appendix.
\begin{enumerate}[label=A.\arabic*]
\item
  \textbf{Symbols and Equations}
  In Section~\ref{sxn:appendix_A}, we summarize the important symbols and key results,
         including the dimensions of the vectors and matrices, the different notations for the energies,
         and the key equations necessary here and later.
   \item
   \textbf{Summary of the \SMOG}
         In Section~\ref{sxn:summary_sst92}, we provide a more detailed analysis of the results we derive in Section~\ref{sxn:SMOG_main-st_av}.
         In particular, in Section~\ref{app:st-gen-err-annealed-ham}, we repeat the derivations of the ST \GeneralizationError $\AVGSTGE$ and related quantities (from Section~\ref{sxn:mathP}),
         using more concrete notation to align with \cite{SST90, SST92}; and
         in Section~\ref{sxn:appendix_Gan}, we use this to derive the matrix-generalization of the ST \EffectivePotential $\EPSLR$ (as well as the normalization for the weight matrices, necessary for later).
\end{enumerate}

\input{sections/410_smog_traditional}
\input{sections/420_smog_mathP}
\input{sections/430_smog_st_model}
\input{sections/440_smog_gen_gap}
%\input{sections/450_smog_bad_data}
\input{sections/450_new}
\input{sections/460_smog_theory_st_avg}
