
\subsection{Generalization Gap vs. Model Quality}
\label{sxn:SMOG_main-model_quality}

In this subsection, we should distinguish between what is typically done in the \SLT literature versus the \STATMECH approaches.
In \SLT, one is typically interested in modeling the \emph{\GeneralizationGap}.
The \GeneralizationGap quantifies the difference between a models performance on training data versus unseen test data:
\begin{align}
  \label{eqn:gen_gap}
  \mathcal{E}^{emp}_{gap}:= \TTE[\DXtrain]- \TGE[\DXtest]
\end{align}
In contrast, in \STATMECH approaches, one considers the \ModelGeneralizationError directly,
which is sometimes called the \ModelQuality in the \SLT literature.
\michael{I think SST looks at generalization error also.  I think this dichotomy isn't quite correct, and it convovles several issues.  It seems most connected with the par around \EQN~\ref{eqn:Qdef} and \EQN~\ref{eqn:GammaQdef}. Maybe we should combine this with that and put is somewhere, removing the incorrect claim/suggestion, and highlighting the ideas in the next par, which are key.
}
\charles{Thats not what this is about.  This is an important section that belongs here.
needs some work}
Model quality is an indication of the models accuracy, precision, recall, or any other relevant metric based on the task at hand.
%Here, we mean that the \ModelGeneralizationError is a measure of the \ModelQuality on such OOD data.

While related, in developing an analytic theory, the \GeneralizationGap and
the \ModelQuality (or \ModelGeneralizationError) require conceptually different approaches.
This is because the  \GeneralizationGap depends on a specific realization of the training data,
whereas our \ModelGeneralizationError will be formulated on a random training data set
(and then corrected later with empirical data).
In this sense, any theory of the \GeneralizationGap  requires a formalism where the
predicted model error is \Quenched to the training data, which is not what we want.
In contrast, the \ModelGeneralizationError  will be formulated using the \AnnealedApproximation (AA),
and is therefore both conceptually and mathematically simpler.
\michael{These ideas seem key; they should either be combined with the par around \EQN~\ref{eqn:Qdef} and \EQN~\ref{eqn:GammaQdef}, or they should be in the intro.}

\nred{Comment on our paper with YQ}

