\subsection{The Inverse-MP (IMP) Model}
\label{sxn:IMP}
In this section, we rederive the integral $G(\lambda)[IMP]$ for the \InverseMP (IMP) model, focusing on the branch cut starting at $z = \kappa/2$ and extending to infinity. 
This branch cut corresponds to the support of the ESD in this region. 
We will:
\begin{enumerate}
\item Explain the presence of the branch cut and its implications.
\item Show that $R(z)[IMP]$ becomes complex along this branch cut because the term under the square root becomes negative.
\item Perform the integral $G(\lambda)[IMP]$, showing all steps.
\item Compute the \emph{Real} part  $\Re[G(\lambda)[IMP]]$ 
\end{enumerate}

\subsubsection{The Branch Cut in the IMP Model}

The R-transform for the IMP model is given by:
\begin{align}
\label{eqn:iw_r_transf}
R(z)[IMP] = \frac{\kappa - \sqrt{\kappa(\kappa - 2z)}}{z},
\end{align}
where $\kappa > 0$ is a parameter related to the dimensions of the random matrices under consideration.
\michael{Ref.}
The function $\sqrt{\kappa(\kappa - 2z)}$ introduces a branch point at $z = \kappa/2$ because the argument of the square root becomes zero at this point:
\begin{align}
\kappa - 2z = 0 \quad \Rightarrow \quad z = \frac{\kappa}{2}.
\end{align}
For $z > \kappa/2$, the argument $\kappa - 2z$ becomes negative, and thus the square root becomes imaginary. 
This leads to a branch cut starting at $z = \kappa/2$ and extending to $z = \infty$ along the real axis. 
This branch cut affects the analyticity of $R(z)[IMP]$, and it must be carefully considered in the integral $G(\lambda)[IMP]$.

\subsubsection{$R(z)[IMP]$ is Complex Along the Branch Cut}

For $z > \kappa/2$, we have:
\begin{align}
\kappa - 2z < 0 \quad \Rightarrow \quad \sqrt{\kappa(\kappa - 2z)} = \sqrt{-\kappa(2z - \kappa)} = i \sqrt{\kappa(2z - \kappa)}.
\end{align}
%
Therefore, $R(z)[IMP]$ becomes complex:
\begin{align}
R(z)[IMP] = \frac{\kappa - i \sqrt{\kappa(2z - \kappa)}}{z} = \frac{\kappa}{z} - i \frac{ \sqrt{ \kappa(2z - \kappa) } }{ z }.
\end{align}
%
This expression shows that $R(z)[IMP]$ has both real and imaginary parts when $z > \kappa/2$.

\subsubsection{Calculation of $G(\lambda)[IMP]$}

We aim to compute the integral:
\begin{align}
G(\lambda)[IMP] := \int_{z_0}^\lambda \Re[R(z)[IMP]] , dz,
\end{align}
where $z_0 \geq \kappa/2$.

Where the Real part of $R(z)[IMP]$ is
\begin{align}
\text{Re}[R(z)[IMP]] = \frac{\kappa}{z}.
\end{align}
%
The integral of this is
\begin{align}
\Re[G(\lambda)[IMP]] = \int_{z_0}^\lambda \frac{\kappa}{z} , dz = \kappa \left[ \ln z \right]_{z_0}^\lambda = \kappa \left( \ln \lambda - \ln z_0 \right).
\end{align}




